### 2012-08-30 talk/ifl2012.tex
draft of talk

\documentclass{beamer}
% 20+5 minutes
\input{macros}

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

\title{Agda Meets Accelerate}
\date{1.9.2012}
\author{Peter Thiemann \and Manuel Chakravarty}
\institute{%
  University of Freiburg, Germany,\\
  \texttt{thiemann@informatik.uni-freiburg.de}
\and
University of New South Wales, Sydney, Australia,\\
\texttt{chak@cse.unsw.edu.au}
}
\begin{document}

% begin titlepage
% "`normal"' front page template:
%\setbeamertemplate{title page}[default][rounded=true]
\setbeamertemplate{navigation symbols}{}
\begin{frame}[plain,label=fp]
		\advance\textwidth by 3.2em\relax
		\begin{minipage}{\textwidth}\par%
				\maketitle
		\end{minipage}
		\hspace*{2.5em}%
\end{frame}
\makeatother 
% end titlepage
\setbeamertemplate{navigation symbols}[horizontal]

\section{Introduction}

\begin{frame}
  \frametitle{Introduction}
  \begin{itemize}
  \item Embedded languages
  \item Dependent types
  \item Generative programming
  \item Data-parallel programming
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Overview}
  \begin{block}{Mission}
    Construct a generative, data parallel programming language embedded in Agda
  \end{block}
  \begin{block}<2->{Means}
    \dots\ by creating a --- more expressive --- Agda frontend for the
    Accelerate language.
  \end{block}
\end{frame}

\section{Accelerate}
\begin{frame}
  \frametitle{The Accelerate Language}
  \begin{itemize}
  \item Data-parallel array language
  \item Expressive embedding in Haskell
  \item Generates GPU code at run time
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Accelerate Example}
  \framesubtitle{Dot Product}
\begin{verbatim}
dotp :: Vector Float -> Vector Float -> Acc (Scalar Float)
dotp xs ys = let xs' = use xs
                 ys' = use ys
             in  fold (+) 0 (zipWith (*) xs' ys')
\end{verbatim}
  \begin{itemize}
  \item \texttt{dotp xs ys} returns a representation of code that
    computes the dot product of \texttt{xs} and \texttt{ys}
  \item References to inputs via \texttt{use}
  \item Specification of operations via HOAS (viz.\ \texttt{(+)} and \texttt{(*)})
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Running Accelerate Code}
  \begin{itemize}
  \item To execute \texttt{dotp}, Accelerate provides several
    implementations of a \texttt{run} function:
\begin{verbatim}
CUDA.run :: Arrays a => Acc a -> a
\end{verbatim}
  \item This implementation
    \begin{itemize}
    \item compiles the code to GPU kernels
    \item loads the \texttt{use}d arguments along with the code onto
      the GPU
    \item executes and
    \item retrieves the result.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Types in Accelerate}
  \begin{itemize}
  \item \texttt{Vector} and \texttt{Scalar} are suitable
    instantiations of a parameterized array type:
\begin{verbatim}
type Scalar e = Array DIM0 e
type Vector e = Array DIM1 e
\end{verbatim}
  \item Dimensions in turn are abbreviations for certain array shapes:
\begin{verbatim}
type DIM0 = Z
type DIM1 = DIM0 :. Int
-- <and so on>
\end{verbatim}
  \item and simple shapes are built from two basic constructors
\begin{verbatim}
data Z       = Z
data sh :. i = sh :. i
\end{verbatim}
  \end{itemize}
\end{frame}

\section{Agda}
\begin{frame}
  \frametitle{Agda}
  \begin{itemize}
  \item A dependently-typed functional programming language
    \begin{itemize}
    \item Inductive datatype families with dependent pattern matching
    \item Dependent records and parameterized modules
    \item Universe polymorphism
    \item \dots
    \end{itemize}
  \item Interactive proof assistant based on intuitionistic type theory
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Indexed Types}
  \begin{itemize}
  \item Standard example: vector type indexed by its size
\begin{verbatim}
data Vec (A : Set) : Nat -> Set where
  []   : Vec A zero
  _::_ : {n : Nat} -> A -> Vec A n -> Vec A (suc n)
\end{verbatim}
  \item Enables the construction of statically safe access operations
    for vectors
\begin{verbatim}
get : forall {A : Set} {n : Nat} ->
      Vec A n -> (m : Nat) -> m < n -> A
get []        _       ()      -- impossible case
get (x :: xs) zero    p       = x
get (x :: xs) (suc m) (s<s p) = get xs m p
\end{verbatim}
  \end{itemize}
\end{frame}

\section{Dependent Types for Accelerate}

\begin{frame}
  \frametitle{Dependent Types for Accelerate}
  \begin{block}{Expectations}
    \begin{enumerate}
    \item Augment array shapes with static size information
    \item Correctness guarantees for fold-like operations
    \item Express all Haskell examples with stronger guarantees
    \end{enumerate}
  \end{block}
  \begin{block}<2->{Results}
    \begin{enumerate}
    \item Bounds checking works out.
    \item Not satisfactory.
    \item Some workarounds required.
    \end{enumerate}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Example: Dot Product in Agda-Accelerate}
  \vspace{-\baselineskip}
\begin{verbatim}
dotp : forall {E : Element} {{p : IsNumeric E}} {n : Nat}
     -> PreVector n E -> PreVector n E -> Scalar E
dotp{E} xs ys = 
  let xs' = use xs
      ys' = use ys
  in
  fold _+_ ("0" ::: E) (zipWith _*_ xs' ys')
\end{verbatim}
  \begin{itemize}
  \item \texttt{PreVector n E} : a vector of $n$ elements of
    type \texttt{E}
  \item \texttt{Scalar E} : code that evaluates to a scalar of type \texttt{E}
  \item \texttt{IsNumeric E} restricts \texttt{E} to numeric types
  \end{itemize}
\end{frame}
\begin{frame}[fragile]
  \frametitle{The Array Type}
  \begin{itemize}
  \item The array type is indexed over a shape
  \item Shapes contain size information
\begin{verbatim}
data Shape : Set where
  Z     : Shape
  _:<_> : Shape -> Nat -> Shape
\end{verbatim}
  \item \texttt{Vector} and \texttt{Scalar} are, again,
    specializations of a shape-indexed array type
\begin{verbatim}
Vector n E    = Array    (Z :< n >) E
Scalar E      = Array    Z          E
\end{verbatim}
  \end{itemize}
\end{frame}
\begin{frame}[fragile]
  \frametitle{Bounds Checking: zipWith}
\begin{verbatim}
zipWith : forall {A B C} {sh} -> (Exp A -> Exp B -> Exp C)
        -> Array sh A -> Array sh B -> Array sh C
\end{verbatim}
  \begin{itemize}
  \item Ensures that only arrays of the exact same shape and size are
    zipped together
  \item More restrictive than Accelerate that only checks the
    dimension and then takes the minimum
  \item But that can also be encoded
  \end{itemize}
\end{frame}
\begin{frame}[fragile]
  \frametitle{Accelerate's zipWith Function in Agda}
\begin{verbatim}
zipWith' : forall {A B C} {sha shb} {p : EqShape sha shb}
        -> (Exp A -> Exp B -> Exp C)
        -> Array sha A -> Array shb B
        -> Array (shmin sha shb {p}) C
\end{verbatim}
where\scriptsize
\begin{verbatim}
data EqShape : Shape -> Shape -> Set where
  EqZ : EqShape Z Z
  EqS : {s1 s2 n1 n2} -> EqShape s1 s2 -> EqShape (s1 :< n1 >) (s2 :< n2 >)

shmin : (s1 : Shape) -> (s2 : Shape) -> { p : EqShape s1 s2 } -> Shape
shmin .Z .Z {EqZ} = Z
shmin .(s1 :< n1 >) .(s2 :< n2 >) {EqS {s1} {s2} {n1} {n2} y} =
    shmin s1 s2 {y} :< min n1 n2 >
\end{verbatim}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Bounds Checking: fold and fold1}

  \begin{itemize}
  \item \texttt{fold} and \texttt{fold1} take a binary operation and
    reduce one dimension of an array using the operation
  \item \texttt{fold} works for any array, but \texttt{fold1} requires
    a non-empty array argument
  \item The latter restriction is not expressed in Haskell-Accelerate
  \end{itemize}
  \begin{block}<2->{\vspace{-\baselineskip}}\small
\begin{verbatim}
fold  :: (Shape ix, Elt a) =>
         (Exp a -> Exp a -> Exp a) -> Exp a ->
         Acc (Array (ix :. Int) a) -> Acc (Array ix a)
fold1 :: (Shape ix, Elt a) =>
         (Exp a -> Exp a -> Exp a) ->
         Acc (Array (ix :. Int) a) -> Acc (Array ix a)
\end{verbatim}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Agda Signatures for fold and fold1}
\begin{verbatim}
fold : forall {E}{sh}{n}
     -> (Exp E -> Exp E -> Exp E) -> Exp E
     -> Array (sh :< n >) E 
     -> Array sh E

fold1 : forall {E}{sh}{n}
      -> (Exp E -> Exp E -> Exp E)
      -> Array (sh :< suc n >) E
      -> Array sh E
\end{verbatim}
\end{frame}
\begin{frame}[fragile]
  \frametitle{Another Issue with fold and fold1}
  \begin{itemize}
  \item \texttt{fold f e} and \texttt{fold1 f}  partition their array
    argument and apply \texttt{f} in a bracketing that is a-priori unknown
  \item For deterministic results, \texttt{f} must be associative and
    (in case of \texttt{fold}) \texttt{e} must be a unit of \texttt{f}
  \item<2-> Such restrictions can be expressed in Agda
  \item<3-> Might ask for a signature like this
\begin{verbatim}
fold : ... -> Monoid E
     -> Array (sh :< n >) E -> Array sh E
\end{verbatim}
  \end{itemize}
\end{frame}
\begin{frame}[fragile]
  \frametitle{A Monoidal Catch}
An encoding of monoids in Agda:\scriptsize
\begin{verbatim}
record Monoid (E : Set) : Set where
  field
    op : E -> E -> E
    un : E
    op_assoc : forall {x y z} -> op (op x y) z = op x (op y z)
    op_un_le : forall {x} -> op un x = x
    op_un_re : forall {x} -> op x un = x
\end{verbatim}
\normalsize
\begin{block}<2->{The Catch}
  In the signature of \texttt{fold} \dots
  \begin{itemize}
  \item \texttt{E} is \textbf{not} a \texttt{Set}, but an element of a
    data type that encodes Haskell datatypes
  \item the type of the operation is
    \texttt{(Exp E -> Exp E -> Exp E)}\\
    a HOAS encoding of the operation's \textbf{syntax}
  \item<3-> Associativity of syntax is meaningless \dots 
  \end{itemize}

\end{block}
\end{frame}
\end{document}